Home Nighttime Cough Monitoring

### 2024-10-24 Lilah Melzer/cough_monitor_analysis.qmd
updates for lilah

```{r setup}
#| include: false
library(Hmisc)
library(tidyverse)
library(lubridate)
library(arsenal)
library(rspiro)
library(lmerTest)
library(performance)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Home Nighttime Cough Monitoring",
  "Windows" = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Lilah Melzer/Home Nighttime Cough Monitoring",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Lilah Melzer/Home Nighttime Cough Monitoring"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/GitHub",
  "Windows" = "C:/Users/timvigers/Documents/GitHub",
  "Linux" = "/home/timvigers/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
#| message: false
#| warning: false
# Import REDCap data
source(paste0(github_dir, "/CF/Lilah Melzer/cough_monitor_import.r"))
# Import use data
use <- read.csv("./Data_Clean/Cough study vacation days consent date.csv",
  na.strings = ""
)
# Import counts
agg_counts <- read.csv(
  "./Data_Raw/Cough_w_TotalCounts_0824 - Cough_agg_w_count_0828.csv",
  na.strings = c("", "NMD"), header = F
)
# Format aggregated count data
agg_counts <- data.frame(t(agg_counts))
colnames(agg_counts) <- agg_counts[1, ]
agg_counts$`Date/Device` <- NULL
agg_counts <- agg_counts[-1, ]
colnames(agg_counts)[1] <- "sid"
agg_counts$sid <- sub("FN", "", agg_counts$sid)
# Wide to long, fix dates
agg_counts <- agg_counts %>%
  pivot_longer(
    cols = -one_of("sid"), names_to = "Date", values_to = "Coughs"
  ) %>%
  mutate(
    Date = mdy(Date),
    Coughs = as.numeric(Coughs)
  )
# Add start date
agg_counts$sid <- as.numeric(agg_counts$sid)
agg_counts$Start <-
  mdy(use$Start.date[match(agg_counts$sid, use$Participant.number)])
# Count number of nights
agg_counts <- agg_counts %>%
  filter(Date >= Start) %>%
  group_by(sid) %>%
  mutate(Night = row_number()) %>%
  ungroup()
# Demographics data
data$Diagnosis <- factor(sub(".*arm_*", "", data$redcap_event_name),
  levels = 1:2,
  labels = c("CF", "Non-CF")
)
demo <- data %>%
  group_by(sid) %>%
  filter(row_number() == 1) %>%
  select(
    sid, Diagnosis, age_at_consent, height, weight, pre_fev,
    nih_sex.factor:nih_ethnicity.factor, mutation1.factor, mutation2.factor,
    cftr_mod.factor, cftr_mod_type.factor, mucolytic.factor:antih.factor
  )
# Calculate FEV1% - rspiro not accepting NA values for gender, which is really
# annoying
fev1 <- demo %>%
  select(age_at_consent, height, nih_sex.factor, pre_fev) %>%
  filter(nih_sex.factor %in% c("Male", "Female"))
fev1$pre_fev_pp <- pctpred_GLIgl(fev1$age_at_consent, fev1$height / 100,
  factor(fev1$nih_sex.factor, levels = c("Male", "Female")),
  FEV1 = fev1$pre_fev
)
demo <- left_join(demo, fev1[, c("sid", "pre_fev_pp")])
label(demo$pre_fev_pp) <- "Pre-bronchodilator FEV1 % Predicted"
# CF genotype
demo$Genotype <-
  factor(rowSums(demo[, c("mutation1.factor", "mutation2.factor")] == "F508del"),
    levels = c(2, 1, 0),
    labels = c("F508del homozygous", "F508del heterozygous", "No F508del")
  )
demo[, grep("mutation", colnames(demo))] <- NULL
# Add group to counts
agg_counts$Diagnosis <-
  demo$Diagnosis[match(agg_counts$sid, as.numeric(as.character(demo$sid)))]
# Per night
per_night <- agg_counts %>%
  group_by(sid) %>%
  summarise(
    median_coughs = median(Coughs, na.rm = T),
    nights_data = sum(!is.na(Coughs))
  )
demo$sid <- as.numeric(demo$sid)
demo <- left_join(demo, per_night)
label(demo$median_coughs) <- "Median Coughs per Night"
label(demo$nights_data) <- "Nights With Cough Data"
# Set table options for the document
mycontrols <-
  tableby.control(
    numeric.stats = c("Nmiss", "medianq1q3", "range"),
    numeric.test = "kwt"
  )
```

>>>>

```{r setup}
#| include: false
library(Hmisc)
library(tidyverse)
library(lubridate)
library(arsenal)
library(rspiro)
library(childsds)
library(lmerTest)
library(performance)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Home Nighttime Cough Monitoring",
  "Windows" = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Lilah Melzer/Home Nighttime Cough Monitoring",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Lilah Melzer/Home Nighttime Cough Monitoring"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/GitHub",
  "Windows" = "C:/Users/timvigers/Documents/GitHub",
  "Linux" = "/home/timvigers/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
#| message: false
#| warning: false
# Import REDCap data
source(paste0(github_dir, "/CF/Lilah Melzer/cough_monitor_import.r"))
# Import use data
start_end <- read.csv("./Data_Clean/Start and end dates.csv", na.strings = "")
start_end$Start <- mdy(start_end$Start)
start_end$End <- mdy(start_end$End)
vacations <- read.csv("./Data_Clean/Vacation days.csv", na.strings = "")
vacations$Vacation.Start <- mdy(vacations$Vacation.Start)
vacations$Vacation.End <- mdy(vacations$Vacation.End)
# Import counts
agg_counts <- read.csv(
  "./Data_Raw/Cough_w_TotalCounts_0824 - Cough_agg_w_count_0828.csv",
  na.strings = c("", "NMD"), header = F
)
# Format aggregated count data
agg_counts <- data.frame(t(agg_counts))
colnames(agg_counts) <- agg_counts[1, ]
agg_counts$`Date/Device` <- NULL
agg_counts <- agg_counts[-1, ]
colnames(agg_counts)[1] <- "sid"
agg_counts$sid <- sub("FN", "", agg_counts$sid)
# Wide to long, fix dates
agg_counts <- agg_counts %>%
  pivot_longer(
    cols = -one_of("sid"), names_to = "Date", values_to = "Coughs"
  ) %>%
  mutate(
    Date = mdy(Date),
    Coughs = as.numeric(Coughs)
  )
# Add start and end dates, filter
agg_counts$sid <- as.numeric(agg_counts$sid)
agg_counts <- left_join(agg_counts, start_end)
agg_counts <- agg_counts %>%
  filter(Date >= Start & Date <= End)
# Vacations
agg_counts$Vacation <- apply(agg_counts, 1, function(r) {
  id <- as.numeric(r["sid"])
  d <- r["Date"]
  any(d >= vacations$Vacation.Start[vacations$sid == id] &
    d <= vacations$Vacation.End[vacations$sid == id])
})
agg_counts <- agg_counts %>%
  filter(!Vacation) %>%
  arrange(sid, Date)
# Sick
sick <- data %>%
  select(sid, pulm_ex_v2.factor) %>%
  mutate(
    antib_dt = mdy(antib_dt), 
    px_start = antib_dt - 7,
    px_end = antib_dt + 28
  )
# Count number of nights
agg_counts <- agg_counts %>%
  group_by(sid) %>%
  mutate(Night = row_number()) %>%
  ungroup()
# Demographics data
data$Diagnosis <- factor(sub(".*arm_*", "", data$redcap_event_name),
  levels = 1:2,
  labels = c("CF", "Non-CF")
)
demo <- data %>%
  group_by(sid) %>%
  filter(row_number() == 1) %>%
  select(
    sid, Diagnosis, age_at_consent, height, weight, pre_fev,
    nih_sex.factor:nih_ethnicity.factor, mutation1.factor, mutation2.factor,
    cftr_mod.factor, cftr_mod_type.factor, mucolytic.factor:antih.factor
  )
# Calculate FEV1% - rspiro not accepting NA values for gender, which is really
# annoying
fev1 <- demo %>%
  select(age_at_consent, height, weight, nih_sex.factor, pre_fev) %>%
  filter(nih_sex.factor %in% c("Male", "Female"))
fev1$pre_fev_pp <- pctpred_GLIgl(fev1$age_at_consent, fev1$height / 100,
  factor(fev1$nih_sex.factor, levels = c("Male", "Female")),
  FEV1 = fev1$pre_fev
)
fev1$height_perc <- sds(as.numeric(fev1$height),
  age = fev1$age_at_consent, sex = fev1$nih_sex.factor,
  male = "Male", female = "Female", item = "height2_20", ref = cdc.ref,
  type = "perc"
) * 100
fev1$weight_perc <- sds(as.numeric(fev1$weight),
  age = fev1$age_at_consent, sex = fev1$nih_sex.factor,
  male = "Male", female = "Female", item = "weight2_20", ref = cdc.ref,
  type = "perc"
) * 100
demo <-
  left_join(demo, fev1[, c("sid", "height_perc", "weight_perc", "pre_fev_pp")])
label(demo$pre_fev_pp) <- "Pre-bronchodilator FEV1 % Predicted"
label(demo$height_perc) <- "Height %ile"
label(demo$weight_perc) <- "Weight %ile"
# CF genotype
demo$Genotype <-
  factor(rowSums(demo[, c("mutation1.factor", "mutation2.factor")] == "F508del"),
    levels = c(2, 1, 0),
    labels = c("F508del homozygous", "F508del heterozygous", "No F508del")
  )
demo[, grep("mutation", colnames(demo))] <- NULL
# Add group to counts
agg_counts$Diagnosis <-
  demo$Diagnosis[match(agg_counts$sid, as.numeric(as.character(demo$sid)))]
# Per night
per_night <- agg_counts %>%
  group_by(sid) %>%
  summarise(
    median_coughs = median(Coughs, na.rm = T),
    nights_data = sum(!is.na(Coughs))
  )
demo$sid <- as.numeric(demo$sid)
demo <- left_join(demo, per_night)
label(demo$median_coughs) <- "Median Coughs per Night"
label(demo$nights_data) <- "Nights With Cough Data"
# Order columns
demo <- demo %>%
  select(sid, height_perc, weight_perc, pre_fev_pp, everything()) %>%
  select(-height, -weight, -pre_fev)
# Exclude vacation days

# Set table options for the document
mycontrols <-
  tableby.control(
    numeric.stats = c("Nmiss", "medianq1q3", "range"),
    numeric.test = "kwt"
  )
```
